---
title: "ISYE 8803: Modeling the Cost Effectiveness of Biologics for Systemic Lupus Erythematosus in Black Women in the United States: A Markov Analysis"
author: "Taofik Ahmed Suleiman and Elizabeth Osota"
date: "April 20, 2025"
output:
  html_document:
    df_print: paged
  pdf_document:
    keep_tex: true
always_allow_html: true
---
## Helper Functions
```{r}
source('../helper_functions/load_packages.R') ## goes up a directory and then into the folder helper_functions.
source("../helper_functions/STM Functions.R")
source("../helper_functions/MarkovModelFunctions.R")

OR_to_RR <- function(OR, base_prob) {
  if (OR <= 0) {
    stop("Odds ratio must be greater than 0")
  }
  RR <- OR / (1 - base_prob + (base_prob * OR))
  
  return(RR)
}
```

## Libraries

```{r}
library(knitr)
library(ggplot2)
library(DiagrammeR)
library(DiagrammeRsvg)
library(rsvg)
library(dplyr)
library(kableExtra)
library(tidyverse)
library(scales)
library(reshape2)
library(gridExtra)
library(readxl)
library(dampack)
```

## Reading in the parameters that describe our model

```{r}
source('Final_Project_Inputs_Apr20.R')
params <- input
```

```{r}
## Process Age-Specific Mortality Data

df_mortality_rates_by_age <- read_excel("2021_female_black_life_table.xlsx", sheet = "Table 15")

v_r_RD_by_age <-  as.matrix(df_mortality_rates_by_age$`Mortality rate age x`)
names(v_r_RD_by_age) <- df_mortality_rates_by_age$Age

# Convert annual mortality rates to 6-month cycle length
cycle_adjustment <- params$cycle_length / 12
v_r_mort_by_age_adj <- 1 - (1 - v_r_RD_by_age)^cycle_adjustment

# Extract age-specific all-cause mortality for ages in model time horizon
# Vector of mortality rates for each cycle in our model
age_indices <- params$mean_age + 0:(params$num_cycles - 1) * cycle_adjustment
age_indices <- ceiling(age_indices)  
max_age <- max(as.numeric(names(v_r_RD_by_age)))
age_indices[age_indices > max_age] <- max_age

v_r_R_D_cycle <- numeric(length(age_indices))

for (i in 1:length(age_indices)) {
  age_index <- min(age_indices[i] + 1, length(v_r_RD_by_age))
  v_r_R_D_cycle[i] <- v_r_RD_by_age[age_index]
}

# Calculate mortality rates for each health state by cycle
v_r_M_D_cycle <- v_r_R_D_cycle * params$hr_M_D
v_r_S_D_cycle <- v_r_R_D_cycle * params$hr_S_D
v_r_OD_D_cycle <- v_r_R_D_cycle * params$hr_OD_D

# Convert rates to probabilities for each cycle
v_p_R_D_cycle <- rate_to_prob(v_r_R_D_cycle)
v_p_M_D_cycle <- rate_to_prob(v_r_M_D_cycle)
v_p_S_D_cycle <- rate_to_prob(v_r_S_D_cycle)
v_p_OD_D_cycle <- rate_to_prob(v_r_OD_D_cycle)
```

## Define Health States and Transitions

```{r}
# Define new states for the expanded model
states <- c("Remission", "Active_Moderate", "Active_Severe", "Organ_Damage", "Death")
v_names_states <- c("R", "M", "S", "OD", "D")
n_states <- length(v_names_states)
n_cycles <- params$num_cycles

# Define strategies
v_names_str <- c("SoC", "Belimumab + SoC", "Anifrolumab + SoC") 
n_str <- length(v_names_str)

# Vector of discount factors for all cycles
v_discount <- 1 / ((1 + params$d_c) ^ ((0:params$num_cycles) * params$cycle_length))
```


## Standard Care Transition Matrix

```{r}

# Initialize array
trans_mat_standard <- array(0, 
        dim = c(n_states, n_states, n_cycles),
        dimnames = list(v_names_states, v_names_states, 1:(n_cycles)))

for(t in 1:n_cycles) {
  # From Remission
  trans_mat_standard["R", "R", t] <- (1 - v_p_R_D_cycle[t]) * params$p_R_R
  trans_mat_standard["R", "M", t] <- (1 - v_p_R_D_cycle[t]) * params$p_R_M
  trans_mat_standard["R", "S", t] <- (1 - v_p_R_D_cycle[t]) * params$p_R_S
  trans_mat_standard["R", "OD", t] <- (1 - v_p_R_D_cycle[t]) * params$p_R_OD
  trans_mat_standard["R", "D", t] <- v_p_R_D_cycle[t]
  
  # From Active Moderate
  trans_mat_standard["M", "M", t] <- (1 - v_p_M_D_cycle[t]) * params$p_M_M
  trans_mat_standard["M", "R", t] <- (1 - v_p_M_D_cycle[t]) * params$p_M_R
  trans_mat_standard["M", "OD", t] <- (1 - v_p_M_D_cycle[t]) * params$p_M_OD
  trans_mat_standard["M", "S", t] <- (1 - v_p_M_D_cycle[t]) * params$p_M_S
  trans_mat_standard["M", "D", t] <- v_p_M_D_cycle[t]
  
  # From Active Severe
  trans_mat_standard["S", "S", t] <- (1 - v_p_S_D_cycle[t]) * params$p_S_S
  trans_mat_standard["S", "M", t] <- (1 - v_p_S_D_cycle[t]) * params$p_S_M
  trans_mat_standard["S", "R", t] <- (1 - v_p_S_D_cycle[t]) * params$p_S_R
  trans_mat_standard["S", "OD", t] <- (1 - v_p_S_D_cycle[t]) * params$p_S_OD
  trans_mat_standard["S", "D", t] <- v_p_S_D_cycle[t]
  
  # From Organ Damage
  trans_mat_standard["OD", "OD", t] <- (1 - v_p_OD_D_cycle[t]) * params$p_OD_OD
  trans_mat_standard["OD", "D", t] <- v_p_OD_D_cycle[t]
  
  # From Death
  trans_mat_standard["D", "D", t] <- 1
}

check_sum_of_transition_array(trans_mat_standard, n_states = n_states, n_cycles = n_cycles, verbose = TRUE)

```


## Belimumab Transition Matrix

```{r}
trans_mat_belimumab <- trans_mat_standard

for(t in 1:n_cycles) {
  # SOC transition probabilities
  SOC_p_R_M <- trans_mat_standard["R", "M", t] / (1 - v_p_R_D_cycle[t])
  SOC_p_R_S <- trans_mat_standard["R", "S", t] / (1 - v_p_R_D_cycle[t])
  SOC_p_M_S <- trans_mat_standard["M", "S", t] / (1 - v_p_M_D_cycle[t])
  
  SOC_p_S_M <- trans_mat_standard["S", "M", t] / (1 - v_p_S_D_cycle[t])
  SOC_p_M_R <- trans_mat_standard["M", "R", t] / (1 - v_p_M_D_cycle[t])
  SOC_p_S_R <- trans_mat_standard["S", "R", t] / (1 - v_p_S_D_cycle[t])
  
  SOC_p_R_OD <- trans_mat_standard["R", "OD", t] / (1 - v_p_R_D_cycle[t])
  SOC_p_M_OD <- trans_mat_standard["M", "OD", t] / (1 - v_p_M_D_cycle[t])
  SOC_p_S_OD <- trans_mat_standard["S", "OD", t] / (1 - v_p_S_D_cycle[t])
  
  # Convert odds ratio to relative risk
  rr_S_M <- OR_to_RR(params$or_improvement_bel, SOC_p_S_M)
  rr_M_R <- OR_to_RR(params$or_improvement_bel, SOC_p_M_R)
  rr_S_R <- OR_to_RR(params$or_improvement_bel, SOC_p_S_R)
  
  
  # Update progression transitions
  trans_mat_belimumab["R", "M", t] <- (1 - v_p_R_D_cycle[t]) * rate_to_prob(prob_to_rate(SOC_p_R_M) * params$hr_progression_bel)
  trans_mat_belimumab["R", "S", t] <- (1 - v_p_R_D_cycle[t]) * rate_to_prob(prob_to_rate(SOC_p_R_S) * params$hr_progression_bel)
  trans_mat_belimumab["M", "S", t] <- (1 - v_p_M_D_cycle[t]) * rate_to_prob(prob_to_rate(SOC_p_M_S) * params$hr_progression_bel)

  trans_mat_belimumab["R", "OD", t] <- (1 - v_p_R_D_cycle[t]) * rate_to_prob(prob_to_rate(SOC_p_R_OD) * params$hr_progression_bel)
  trans_mat_belimumab["M", "OD", t] <- (1 - v_p_M_D_cycle[t]) * rate_to_prob(prob_to_rate(SOC_p_M_OD) * params$hr_progression_bel)
  trans_mat_belimumab["S", "OD", t] <- (1 - v_p_S_D_cycle[t]) * rate_to_prob(prob_to_rate(SOC_p_S_OD) * params$hr_progression_bel)
  
  # Update improvement transitions
  trans_mat_belimumab["S", "M", t] <- (1 - v_p_S_D_cycle[t]) * SOC_p_S_M * rr_S_M
  trans_mat_belimumab["M", "R", t] <- (1 - v_p_M_D_cycle[t]) * SOC_p_M_R * rr_M_R
  trans_mat_belimumab["S", "R", t] <- (1 - v_p_S_D_cycle[t]) * SOC_p_S_R * rr_S_R
  
  # Update self-transitions
  trans_mat_belimumab["R", "R", t] <- (1 - v_p_R_D_cycle[t]) - (
    trans_mat_belimumab["R", "M", t] + 
    trans_mat_belimumab["R", "S", t] + 
    trans_mat_belimumab["R", "OD", t]
  )
  
  trans_mat_belimumab["M", "M", t] <- (1 - v_p_M_D_cycle[t]) - (
    trans_mat_belimumab["M", "R", t] + 
    trans_mat_belimumab["M", "S", t] + 
    trans_mat_belimumab["M", "OD", t]
  )
  
  trans_mat_belimumab["S", "S", t] <- (1 - v_p_S_D_cycle[t]) - (
    trans_mat_belimumab["S", "M", t] + 
    trans_mat_belimumab["S", "R", t] + 
    trans_mat_belimumab["S", "OD", t]
  )
}

check_sum_of_transition_array(trans_mat_belimumab, n_states = n_states, n_cycles = n_cycles, verbose = TRUE)
```

## Anifrolumab Transition Matrix

```{r}
trans_mat_anifrolumab <- trans_mat_standard

for(t in 1:n_cycles) {
  # SOC transition probabilities
  SOC_p_R_M <- trans_mat_standard["R", "M", t] / (1 - v_p_R_D_cycle[t])
  SOC_p_R_S <- trans_mat_standard["R", "S", t] / (1 - v_p_R_D_cycle[t])
  SOC_p_M_S <- trans_mat_standard["M", "S", t] / (1 - v_p_M_D_cycle[t])
  
  SOC_p_S_M <- trans_mat_standard["S", "M", t] / (1 - v_p_S_D_cycle[t])
  SOC_p_M_R <- trans_mat_standard["M", "R", t] / (1 - v_p_M_D_cycle[t])
  SOC_p_S_R <- trans_mat_standard["S", "R", t] / (1 - v_p_S_D_cycle[t])
  
  SOC_p_R_OD <- trans_mat_standard["R", "OD", t] / (1 - v_p_R_D_cycle[t])
  SOC_p_M_OD <- trans_mat_standard["M", "OD", t] / (1 - v_p_M_D_cycle[t])
  SOC_p_S_OD <- trans_mat_standard["S", "OD", t] / (1 - v_p_S_D_cycle[t])
  
  # Convert odds ratio to relative risk
  rr_S_M <- OR_to_RR(params$or_improvement_anif, SOC_p_S_M)
  rr_M_R <- OR_to_RR(params$or_improvement_anif, SOC_p_M_R)
  rr_S_R <- OR_to_RR(params$or_improvement_anif, SOC_p_S_R)
  
  # Update progression transitions
  trans_mat_anifrolumab["R", "M", t] <- (1 - v_p_R_D_cycle[t]) * SOC_p_R_M * params$rr_progression_anif
  trans_mat_anifrolumab["R", "S", t] <- (1 - v_p_R_D_cycle[t]) * SOC_p_R_S * params$rr_progression_anif
  trans_mat_anifrolumab["M", "S", t] <- (1 - v_p_M_D_cycle[t]) * SOC_p_M_S * params$rr_progression_anif
  
  trans_mat_anifrolumab["R", "OD", t] <- (1 - v_p_R_D_cycle[t]) * SOC_p_R_OD * params$rr_progression_anif
  trans_mat_anifrolumab["M", "OD", t] <- (1 - v_p_M_D_cycle[t]) * SOC_p_M_OD * params$rr_progression_anif
  trans_mat_anifrolumab["S", "OD", t] <- (1 - v_p_S_D_cycle[t]) * SOC_p_S_OD * params$rr_progression_anif
  
  # Update improvement transitions
  trans_mat_anifrolumab["S", "M", t] <- (1 - v_p_S_D_cycle[t]) * SOC_p_S_M * rr_S_M
  trans_mat_anifrolumab["M", "R", t] <- (1 - v_p_M_D_cycle[t]) * SOC_p_M_R * rr_M_R
  trans_mat_anifrolumab["S", "R", t] <- (1 - v_p_S_D_cycle[t]) * SOC_p_S_R * rr_S_R
  
  # Update self-transitions
  trans_mat_anifrolumab["R", "R", t] <- (1 - v_p_R_D_cycle[t]) - (
    trans_mat_anifrolumab["R", "M", t] + 
    trans_mat_anifrolumab["R", "S", t] + 
    trans_mat_anifrolumab["R", "OD", t]
  )
  
  trans_mat_anifrolumab["M", "M", t] <- (1 - v_p_M_D_cycle[t]) - (
    trans_mat_anifrolumab["M", "R", t] + 
    trans_mat_anifrolumab["M", "S", t] + 
    trans_mat_anifrolumab["M", "OD", t]
  )
  
  trans_mat_anifrolumab["S", "S", t] <- (1 - v_p_S_D_cycle[t]) - (
    trans_mat_anifrolumab["S", "M", t] + 
    trans_mat_anifrolumab["S", "R", t] + 
    trans_mat_anifrolumab["S", "OD", t]
  )
}

check_sum_of_transition_array(trans_mat_anifrolumab, n_states = n_states, n_cycles = n_cycles, verbose = TRUE)

```

# Run Markov Trace with age-dependent transtion
```{r}
sim_standard <- matrix(0, 
                 nrow     = (n_cycles + 1), 
                 ncol = n_states, 
                 dimnames = list(0:n_cycles, v_names_states))
v_s_init <- c(R = 0, M = params$moderate_ratio, S = params$severe_ratio, OD = 0, D = 0)

sim_standard[1, ] <- v_s_init

## Initialize cohort trace for strategies
sim_belimumab <- sim_anifrolumab <- sim_standard # structure and initial states remain the same.

## Iterative solution of age-dependent cSTM
for(t in 1:n_cycles){
  ## Fill in cohort trace for each strategy
  sim_standard[t + 1, ]      <- sim_standard[t, ] %*% trans_mat_standard[, , t] # This computes the alpha_{n+1}=alpha_n*P_n step, where t is indicating the cycle
  sim_belimumab[t + 1, ]      <- sim_belimumab[t, ] %*% trans_mat_belimumab[, , t]
  sim_anifrolumab[t + 1, ]      <- sim_anifrolumab[t, ] %*% trans_mat_anifrolumab[, , t]
}

## Store the cohort traces in a list
l_m_M <- list(sim_standard,
              sim_belimumab,
              sim_anifrolumab)
names(l_m_M) <- v_names_str

```

## Plot the Cohort Traces
```{r}
# Prepare data for plotting
prepare_plot_df <- function(sim_matrix, strategy) {
  df <- data.frame(Cycle = 0:n_cycles, sim_matrix)
  df_melted <- reshape2::melt(df, id.vars = "Cycle", 
                              variable.name = "State", 
                              value.name = "Proportion")
  df_melted$Strategy <- strategy
  df_melted$Age <- params$mean_age + df_melted$Cycle * (params$cycle_length/12)
  return(df_melted)
}

# Combine all strategy data
plot_df <- rbind(
  prepare_plot_df(sim_standard, "Standard of Care"),
  prepare_plot_df(sim_belimumab, "Belimumab + SOC"),
  prepare_plot_df(sim_anifrolumab, "Anifrolumab + SOC")
)

# Generate the plot
state_plot <- ggplot(plot_df, aes(x = Age, y = Proportion, color = State)) +
  geom_line(linewidth = 1.1) +
  facet_wrap(~ Strategy, ncol = 1) +
  theme_minimal(base_size = 14) +
  labs(
    title = "State Distribution Over Time by Strategy",
    y = "Proportion of Patients",
    x = "Age (Years)"
  ) +
  theme(
    legend.position = "bottom",
    plot.title = element_text(hjust = 0.5, face = "bold")
  )

# Save
png("figures/state_distribution_plot.png", width = 1000, height = 800, res = 150)
print(state_plot)
dev.off()

# Or just display it in the document
print(state_plot)

```


## Cost, and Utility

```{r}

# Cost vectors per strategy (in USD per 6-month cycle)
costs_standard <- c(
  R  = params$c_R,  # Lower cost in remission
  M  = params$c_M,       # Active Moderate
  S  = params$c_S,  # Active Severe 
  OD = params$c_OD,        # Organ Damage
  D  = 0  # Death
)

costs_belimumab <- c(
  R  = params$c_R + params$cost_belimumab - params$c_low_dose_steriod_reduction,
  M  = params$c_M + params$cost_belimumab - params$c_low_dose_steriod_reduction,
  S  = params$c_S + params$cost_belimumab,
  OD = params$c_OD,
  D  = 0
)

costs_anifrolumab <- c(
  R  = params$c_R + params$cost_anifrolumab - params$c_low_dose_steriod_reduction,
  M  = params$c_M + params$cost_anifrolumab - params$c_low_dose_steriod_reduction,
  S  = params$c_S + params$cost_anifrolumab,
  OD = params$c_OD,
  D  = 0
)

# Utility values per state
utilities_SoC <- c(
  R  = params$u.remission,       # Remission/Low
  M  = params$u.active,          # Active Moderate
  S  = params$u.severe,          # Active Severe
  OD = params$u.organ_damage,    # Organ Damage
  D  = params$u.death            # Death
)

utilities_trt <- c(
  R  = params$u.remission + params$u_R_M_trt,       # Remission/Low
  M  = params$u.active + params$u_R_M_trt,          # Active Moderate
  S  = params$u.severe,          # Active Severe
  OD = params$u.organ_damage,    # Organ Damage
  D  = params$u.death            # Death
)

# Store costs and utilities in lists
l_costs <- list(costs_standard, costs_belimumab, costs_anifrolumab)
names(l_costs) <- v_names_str

l_u <- list(utilities_SoC, utilities_trt, utilities_trt)
names(l_u) <- v_names_str
```


## Total Cost and QALYs

```{r}

# Function to Calculate Total Cost and QALYs
# Updated list of states (5-state model)
states <- c("R", "M", "S", "OD", "D")
num_states <- length(states)
n_cycles <- params$num_cycles


# Initialize state distribution matrix
initialize_matrix <- function() {
  mat <- matrix(0, nrow = n_cycles + 1, ncol = num_states)
  colnames(mat) <- states
  mat[1, "M"] <- params$moderate_ratio  # 65% of patients start in Moderate state
  mat[1, "S"] <- params$severe_ratio  # 35% of patients start in Severe state
  return(mat)
}
## Within-cycle correction (WCC) using half-cycle correction ----
v_wcc_half <- gen_wcc(n_cycles = n_cycles,  
                 method = "half-cycle") # vector of wcc

# Calculate total costs and QALYs with discounting and half-cycle correction
calculate_totals <- function(sim_matrix, strategy_name, costs, utilities) {
  # Get state values for each cycle
  v_costs_cycle <- numeric(n_cycles + 1)
  v_qalys_cycle <- numeric(n_cycles + 1)
  
  # Calculate costs and QALYs for each cycle
  for (t in 1:(n_cycles + 1)) {
    v_costs_cycle[t] <- sum(sim_matrix[t,] * costs)
    v_qalys_cycle[t] <- sum(sim_matrix[t,] * utilities) * (params$cycle_length/12)
  }
  
  # Apply half-cycle correction and discounting
  total_cost <- sum(v_costs_cycle * v_wcc_half * v_discount[1:(n_cycles+1)])
  total_qaly <- sum(v_qalys_cycle * v_wcc_half * v_discount[1:(n_cycles+1)])
  
  return(data.frame(
    Strategy = strategy_name,
    Total_Cost = total_cost,
    Total_QALY = total_qaly
  ))
}

# Generate Results
results <- rbind(
  calculate_totals(sim_standard, "Standard of Care", costs_standard, utilities_SoC),
  calculate_totals(sim_belimumab, "Belimumab + SOC", costs_belimumab, utilities_trt),
  calculate_totals(sim_anifrolumab, "Anifrolumab + SOC", costs_anifrolumab, utilities_trt)
)

results_melted <- reshape2::melt(results, id.vars = "Strategy",
                                 variable.name = "Measure",
                                 value.name = "Value")

# Plot: Total Cost per Strategy
cost_plot <- ggplot(subset(results_melted, Measure == "Total_Cost"), 
                    aes(x = Strategy, y = Value, fill = Strategy)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal(base_size = 14) +
  labs(title = "Total Costs per Strategy",
       y = "Total Cost ($)", x = "") +
  theme(
    axis.text.x = element_text(angle = 15, hjust = 1),
    legend.position = "none",
    panel.grid = element_blank()
    ) +
  scale_fill_brewer(palette = "Set2")

# Save to file
#png("figures/total_costs_per_strategy.png", width = 1000, height = 800, res = 150)
print(cost_plot)
#dev.off()

# Plot: Total QALYs per Strategy
qaly_plot <- ggplot(subset(results_melted, Measure == "Total_QALY"), 
                    aes(x = Strategy, y = Value, fill = Strategy)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal(base_size = 14) +
  labs(title = "Total QALYs per Strategy",
       y = "Total QALYs", x = "") +
  theme(
    axis.text.x = element_text(angle = 15, hjust = 1),
    legend.position = "none",
    panel.grid = element_blank()
    ) +
  scale_fill_brewer(palette = "Set2")

#ggsave("figures/total_qalys_per_strategy.png", plot = qaly_plot,
       #width = 10, height = 6, dpi = 300)

# Save to file
#png("figures/total_qalys_per_strategy.png", width = 1000, height = 800, res = 150)

print(qaly_plot)
#dev.off()

```

## ICER Table

```{r}

## ICER Table

# Reorder strategies by increasing cost
icer_input <- results[order(results$Total_Cost), ]

# Compute Incremental Cost, QALY, and ICER
icer_input$Inc_Cost <- c(NA, diff(icer_input$Total_Cost))
icer_input$Inc_QALY <- c(NA, diff(icer_input$Total_QALY))
icer_input$ICER <- round(icer_input$Inc_Cost / icer_input$Inc_QALY, 2)

# Round other columns
icer_input <- icer_input %>%
  mutate(across(c(Total_Cost, Total_QALY, Inc_Cost, Inc_QALY), round, 2))

# Display Table
kable(icer_input, caption = "Incremental Cost-Effectiveness Ratios (ICER) Table") %>%
  kable_styling(full_width = FALSE, position = "center")

v_costs <- results$Total_Cost
v_qalys <- results$Total_QALY
v_names_str <- results$Strategy

# Perform cost-effectiveness analysis
df_cea <- calculate_icers(cost = v_costs,
                          effect = v_qalys,
                          strategies = v_names_str)

# View the results
print(df_cea)

# Format the CEA table for better presentation
table_cea <- format_table_cea(df_cea)
print(table_cea)

# Visualize the CE frontier
cea_plot <- plot(df_cea, label = "all", txtsize = 16)
print(cea_plot)

```

##  Model Validation

```{r}
calculate_avg_death_age <- function(sim_matrix) {
  # Calculate new deaths at each cycle
  death_proportions <- sim_matrix[, "D"]
  new_deaths <- c(death_proportions[1], diff(death_proportions))
  
  # Calculate time until death
  cycles_until_death <- 0:(length(new_deaths)-1)
  time_until_death <- cycles_until_death * (params$cycle_length/12)
  
  # Calculate weighted average time until death
  avg_time_until_death <- sum(new_deaths * time_until_death) / death_proportions[length(death_proportions)]
  
  # Average age of death = starting age + average time until death
  avg_age_of_death <- params$mean_age + avg_time_until_death
  
  return(avg_age_of_death)
}

# Average age by strategy
avg_age_SoC <- calculate_avg_death_age(sim_standard)
avg_age_Belimumab <- calculate_avg_death_age(sim_belimumab)
avg_age_Anifrolumab <- calculate_avg_death_age(sim_anifrolumab)

# Create results table
death_age_results <- data.frame(
  Strategy = v_names_str,
  Average_Age_of_Death = c(avg_age_SoC, avg_age_Belimumab, avg_age_Anifrolumab)
)

# Display table
kable(death_age_results, caption = "Average Age of Death by Treatment Strategy") %>%
  kable_styling(full_width = FALSE, position = "center")
```

#ICER 

```{r}
# Calculate cost per QALY for each intervention relative to Standard Care
cost_per_qaly_data <- data.frame(
  Strategy = results$Strategy,
  stringsAsFactors = FALSE
)

# Retrieving Standard of Care results
ref_idx <- which(results$Strategy == "Standard of Care")
ref_cost <- results$Total_Cost[ref_idx]
ref_qaly <- results$Total_QALY[ref_idx]

# Calculate the incremental cost per QALY for each strategy
cost_per_qaly_data$ICER <- NA
for (i in 1:nrow(results)) {
  if (i == ref_idx) {
    cost_per_qaly_data$ICER[i] <- 0
  } else {
    inc_cost <- results$Total_Cost[i] - ref_cost
    inc_qaly <- results$Total_QALY[i] - ref_qaly
    cost_per_qaly_data$ICER[i] <- inc_cost / inc_qaly
  }
}

strategy_colors <- c(
  "Anifrolumab + SOC" = "#66c2a5",  # greenish
  "Belimumab + SOC"   = "#fc8d62",  # orangish
  "Standard of Care"     = "#8da0cb"   # bluish
)

# Create results bar graph
cost_per_qaly_plot <- ggplot(cost_per_qaly_data, aes(x = Strategy, y = ICER, fill = Strategy)) +
  geom_bar(stat = "identity") +
  geom_hline(yintercept = 150000, linetype = "dashed", color = "black") +
  annotate("text", x = 1, y = 150000, label = "ICER Willingness to Pay Threshold: $150,000", 
           hjust = -0.1, vjust = -0.5, color = "black") +
  geom_text(aes(label = ifelse(ICER == 0, "Reference", 
                              paste0("$", format(round(ICER, 0), big.mark = ",")))),
            position = position_dodge(width = 0.9), vjust = -0.5) +
  scale_fill_manual(values = strategy_colors) +
  theme_minimal(base_size = 14) +
  labs(
    title = " Incremental Cost-Effectiveness Ratios",
    x = "",
    y = "Cost per QALY Gained ($)"
  )

# Print the plot
print(cost_per_qaly_plot)
```

## One way sensitivity analysis - SOC and Anifrolumab interventions 
```{r}
## One way sensitivity analysis

# Load parameters
source('Final_Project_Inputs_Apr20.R')

# Convert the input data frame to the list format for params_basecase
params_basecase <- as.list(input[1,])

source('Final_Project_Inputs_Apr20.R')
source_input <- input

# Define parameters for one-way sensitivity analysis
owsa_params_range <- data.frame(
  pars = c("cost_anifrolumab", "c_low_dose_steriod_reduction","cost_death", "c_R", "c_M", "c_S", "c_OD", 
           "u.remission", "u.active","u.severe", "u.organ_damage", "rr_progression_anif","or_improvement_anif", 
           "u_R_M_trt", "hr_M_D", "hr_S_D", "hr_OD_D",
           "p_R_R", "p_R_M", "p_R_S", "p_R_OD", "p_M_M", "p_M_R", "p_M_OD", "p_M_S", 
           "p_S_S", "p_S_M", "p_S_R", "p_S_OD"),
  min = c(22430.46*.50, 644.07*.75, 7327.06*.75, 20239.96*.75, 24317.58*.75, 36857.11*.75, 161209.52*.75, 
          0.705*.90, 0.42*.90, 0.37*.90, 0.28*.90, .67*.90, 2.7*.90, 
          0.03*.90, 3.70*.90, 7.15*.90, 12.94*.90, 
          0.79032*.90, 0.124*.90, 0.085*.90, 0.00068*.75, 0.7706*.90,0.1033*.75, 0.0011*.90, 0.125*.90, 
          0.8223*.90, 0.103*.75, 0.053*.90, 0.0217*.90),
  max = c(22430.46*1.50, 644.07*1.25, 7327.06*1.25, 20239.96*1.25, 24317.58*1.25, 36857.11*1.25, 161209.52*1.25,
          0.705*1.10, 0.42*1.10, 0.37*1.10, 0.28*1.10, .67*1.10, 2.7*1.10, 
          0.03*1.10, 3.70*1.10, 7.15*1.10, 12.94*1.10, 
          0.79032*1.10, 0.124*1.10, 0.085*1.10, 0.00068*1.25, 0.7706*1.10, 0.1033*1.25, 0.0011*1.10, 0.125*1.10, 
          0.8223*1.10, 0.103*1.25, 0.053*1.10, 0.0217*1.10)
)


# Create a function to run the model with different parameter values
run_lupus_model <- function(params_updated, wtp = 150000, verbose = FALSE) {
  
  # Create a list of parameters from the input data frame
  params_list <- as.list(source_input[1,])
  
  # Update parameters with the values from sensitivity analysis
  for (param_name in names(params_updated)) {
    params_list[[param_name]] <- params_updated[[param_name]]
  }
  
  p <- params_list
  
  # Process Age-Specific Mortality Data
  df_mortality_rates_by_age <- read_excel("2021_female_black_life_table.xlsx", sheet = "Table 15")
  
  v_r_RD_by_age <-  as.matrix(df_mortality_rates_by_age$`Mortality rate age x`)
  names(v_r_RD_by_age) <- df_mortality_rates_by_age$Age
  
  # Convert annual mortality rates to 6-month cycle length
  cycle_adjustment <- p$cycle_length / 12
  v_r_mort_by_age_adj <- 1 - (1 - v_r_RD_by_age)^cycle_adjustment
  
  # Extract age-specific all-cause mortality for ages in model time horizon
  age_indices <- p$mean_age + 0:(p$num_cycles - 1) * cycle_adjustment
  age_indices <- ceiling(age_indices)  
  max_age <- max(as.numeric(names(v_r_RD_by_age)))
  age_indices[age_indices > max_age] <- max_age
  
  v_r_R_D_cycle <- numeric(length(age_indices))
  
  for (i in 1:length(age_indices)) {
    age_index <- min(age_indices[i] + 1, length(v_r_RD_by_age))
    v_r_R_D_cycle[i] <- v_r_RD_by_age[age_index]
  }
  
  # Calculate mortality rates for each health state by cycle
  v_r_M_D_cycle <- v_r_R_D_cycle * p$hr_M_D
  v_r_S_D_cycle <- v_r_R_D_cycle * p$hr_S_D
  v_r_OD_D_cycle <- v_r_R_D_cycle * p$hr_OD_D
  
  # Convert rates to probabilities for each cycle
  v_p_R_D_cycle <- rate_to_prob(v_r_R_D_cycle)
  v_p_M_D_cycle <- rate_to_prob(v_r_M_D_cycle)
  v_p_S_D_cycle <- rate_to_prob(v_r_S_D_cycle)
  v_p_OD_D_cycle <- rate_to_prob(v_r_OD_D_cycle)
  
  # Define health states
  states <- c("Remission", "Active_Moderate", "Active_Severe", "Organ_Damage", "Death")
  v_names_states <- c("R", "M", "S", "OD", "D")
  n_states <- length(v_names_states)
  n_cycles <- p$num_cycles
  
  # Define strategies
  v_names_str <- c("SoC", "Anifrolumab + SoC") 
  n_str <- length(v_names_str)
  
  # Vector of discount factors for all cycles
  v_discount <- 1 / ((1 + p$d_c) ^ ((0:p$num_cycles) * p$cycle_length))
  
  # Within-cycle correction (WCC) using half-cycle correction
  v_wcc_half <- gen_wcc(n_cycles = n_cycles, method = "half-cycle")
  
  # Standard Care Transition Matrix
  trans_mat_standard <- array(0, 
                             dim = c(n_states, n_states, n_cycles),
                             dimnames = list(v_names_states, v_names_states, 1:(n_cycles)))
  
  for(t in 1:n_cycles) {
    # From Remission
    trans_mat_standard["R", "R", t] <- (1 - v_p_R_D_cycle[t]) * p$p_R_R
    trans_mat_standard["R", "M", t] <- (1 - v_p_R_D_cycle[t]) * p$p_R_M
    trans_mat_standard["R", "S", t] <- (1 - v_p_R_D_cycle[t]) * p$p_R_S
    trans_mat_standard["R", "OD", t] <- (1 - v_p_R_D_cycle[t]) * p$p_R_OD
    trans_mat_standard["R", "D", t] <- v_p_R_D_cycle[t]
    
    # From Active Moderate
    trans_mat_standard["M", "M", t] <- (1 - v_p_M_D_cycle[t]) * p$p_M_M
    trans_mat_standard["M", "R", t] <- (1 - v_p_M_D_cycle[t]) * p$p_M_R
    trans_mat_standard["M", "OD", t] <- (1 - v_p_M_D_cycle[t]) * p$p_M_OD
    trans_mat_standard["M", "S", t] <- (1 - v_p_M_D_cycle[t]) * p$p_M_S
    trans_mat_standard["M", "D", t] <- v_p_M_D_cycle[t]
    
    # From Active Severe
    trans_mat_standard["S", "S", t] <- (1 - v_p_S_D_cycle[t]) * p$p_S_S
    trans_mat_standard["S", "M", t] <- (1 - v_p_S_D_cycle[t]) * p$p_S_M
    trans_mat_standard["S", "R", t] <- (1 - v_p_S_D_cycle[t]) * p$p_S_R
    trans_mat_standard["S", "OD", t] <- (1 - v_p_S_D_cycle[t]) * p$p_S_OD
    trans_mat_standard["S", "D", t] <- v_p_S_D_cycle[t]
    
    # From Organ Damage
    trans_mat_standard["OD", "OD", t] <- (1 - v_p_OD_D_cycle[t]) * p$p_OD_OD
    trans_mat_standard["OD", "D", t] <- v_p_OD_D_cycle[t]
    
    # From Death
    trans_mat_standard["D", "D", t] <- 1
  }
  
  # Anifrolumab Transition Matrix
  trans_mat_anifrolumab <- trans_mat_standard
  
  for(t in 1:n_cycles) {
    # SOC transition probabilities
    SOC_p_R_M <- trans_mat_standard["R", "M", t] / (1 - v_p_R_D_cycle[t])
    SOC_p_R_S <- trans_mat_standard["R", "S", t] / (1 - v_p_R_D_cycle[t])
    SOC_p_M_S <- trans_mat_standard["M", "S", t] / (1 - v_p_M_D_cycle[t])
    
    SOC_p_S_M <- trans_mat_standard["S", "M", t] / (1 - v_p_S_D_cycle[t])
    SOC_p_M_R <- trans_mat_standard["M", "R", t] / (1 - v_p_M_D_cycle[t])
    SOC_p_S_R <- trans_mat_standard["S", "R", t] / (1 - v_p_S_D_cycle[t])
    
    SOC_p_R_OD <- trans_mat_standard["R", "OD", t] / (1 - v_p_R_D_cycle[t])
    SOC_p_M_OD <- trans_mat_standard["M", "OD", t] / (1 - v_p_M_D_cycle[t])
    SOC_p_S_OD <- trans_mat_standard["S", "OD", t] / (1 - v_p_S_D_cycle[t])
    
    # Convert odds ratio to relative risk
    rr_S_M <- OR_to_RR(p$or_improvement_anif, SOC_p_S_M)
    rr_M_R <- OR_to_RR(p$or_improvement_anif, SOC_p_M_R)
    rr_S_R <- OR_to_RR(p$or_improvement_anif, SOC_p_S_R)
    
    # Update progression transitions
    trans_mat_anifrolumab["R", "M", t] <- (1 - v_p_R_D_cycle[t]) * SOC_p_R_M * p$rr_progression_anif
    trans_mat_anifrolumab["R", "S", t] <- (1 - v_p_R_D_cycle[t]) * SOC_p_R_S * p$rr_progression_anif
    trans_mat_anifrolumab["M", "S", t] <- (1 - v_p_M_D_cycle[t]) * SOC_p_M_S * p$rr_progression_anif
    
    trans_mat_anifrolumab["R", "OD", t] <- (1 - v_p_R_D_cycle[t]) * SOC_p_R_OD * p$rr_progression_anif
    trans_mat_anifrolumab["M", "OD", t] <- (1 - v_p_M_D_cycle[t]) * SOC_p_M_OD * p$rr_progression_anif
    trans_mat_anifrolumab["S", "OD", t] <- (1 - v_p_S_D_cycle[t]) * SOC_p_S_OD * p$rr_progression_anif
    
    # Update improvement transitions
    trans_mat_anifrolumab["S", "M", t] <- (1 - v_p_S_D_cycle[t]) * SOC_p_S_M * rr_S_M
    trans_mat_anifrolumab["M", "R", t] <- (1 - v_p_M_D_cycle[t]) * SOC_p_M_R * rr_M_R
    trans_mat_anifrolumab["S", "R", t] <- (1 - v_p_S_D_cycle[t]) * SOC_p_S_R * rr_S_R
    
    # Update self-transitions
    trans_mat_anifrolumab["R", "R", t] <- (1 - v_p_R_D_cycle[t]) - (
      trans_mat_anifrolumab["R", "M", t] + 
        trans_mat_anifrolumab["R", "S", t] + 
        trans_mat_anifrolumab["R", "OD", t]
    )
    
    trans_mat_anifrolumab["M", "M", t] <- (1 - v_p_M_D_cycle[t]) - (
      trans_mat_anifrolumab["M", "R", t] + 
        trans_mat_anifrolumab["M", "S", t] + 
        trans_mat_anifrolumab["M", "OD", t]
    )
    
    trans_mat_anifrolumab["S", "S", t] <- (1 - v_p_S_D_cycle[t]) - (
      trans_mat_anifrolumab["S", "M", t] + 
        trans_mat_anifrolumab["S", "R", t] + 
        trans_mat_anifrolumab["S", "OD", t]
    )
  }
  
  # Run Markov trace
  sim_standard <- matrix(0, 
                         nrow = (n_cycles + 1), 
                         ncol = n_states, 
                         dimnames = list(0:n_cycles, v_names_states))
  v_s_init <- c(R = 0, M = p$moderate_ratio, S = p$severe_ratio, OD = 0, D = 0)
  
  sim_standard[1, ] <- v_s_init
  
  sim_anifrolumab <- sim_standard
  
  for(t in 1:n_cycles){
    sim_standard[t + 1, ] <- sim_standard[t, ] %*% trans_mat_standard[, , t]
    sim_anifrolumab[t + 1, ] <- sim_anifrolumab[t, ] %*% trans_mat_anifrolumab[, , t]
  }
  
  # Costs and utilities
  costs_standard <- c(
    R = p$c_R,
    M = p$c_M,
    S = p$c_S,
    OD = p$c_OD,
    D = 0
  )
  
  costs_anifrolumab <- c(
    R = p$c_R + p$cost_anifrolumab - p$c_low_dose_steriod_reduction,
    M = p$c_M + p$cost_anifrolumab - p$c_low_dose_steriod_reduction,
    S = p$c_S + p$cost_anifrolumab,
    OD = p$c_OD,
    D = 0
  )
  
  # Utility values per state
  utilities_SoC <- c(
    R = p$u.remission,
    M = p$u.active,
    S = p$u.severe,
    OD = p$u.organ_damage,
    D = p$u.death
  )
  
  utilities_trt <- c(
    R = p$u.remission + p$u_R_M_trt,
    M = p$u.active + p$u_R_M_trt,
    S = p$u.severe,
    OD = p$u.organ_damage,
    D = p$u.death
  )
  
  # Calculate total costs and QALYs with discounting and half-cycle correction
  calculate_totals <- function(sim_matrix, costs, utilities) {
    v_costs_cycle <- numeric(n_cycles + 1)
    v_qalys_cycle <- numeric(n_cycles + 1)
    
    for (t in 1:(n_cycles + 1)) {
      v_costs_cycle[t] <- sum(sim_matrix[t,] * costs)
      v_qalys_cycle[t] <- sum(sim_matrix[t,] * utilities) * (p$cycle_length/12)
    }
    
    total_cost <- sum(v_costs_cycle * v_wcc_half * v_discount[1:(n_cycles+1)])
    total_qaly <- sum(v_qalys_cycle * v_wcc_half * v_discount[1:(n_cycles+1)])
    
    total_nmb <- total_qaly * wtp - total_cost
    
    return(c(Cost = total_cost, QALY = total_qaly, NMB = total_nmb))
  }
  
  # Calculate results for each strategy
  results_SoC <- calculate_totals(sim_standard, costs_standard, utilities_SoC)
  results_Anifrolumab <- calculate_totals(sim_anifrolumab, costs_anifrolumab, utilities_trt)
  
  # Combine results
  df_results <- data.frame(
    Strategy = v_names_str,
    Cost = c(results_SoC["Cost"], results_Anifrolumab["Cost"]),
    QALY = c(results_SoC["QALY"], results_Anifrolumab["QALY"]),
    NMB = c(results_SoC["NMB"], results_Anifrolumab["NMB"])
  )
  
  return(df_results)
}

# Run one-way sensitivity analysis
l_owsa_det <- run_owsa_det(
  params_range = owsa_params_range,
  params_basecase = params_basecase,
  nsamp = 100,
  FUN = run_lupus_model,
  outcomes = c("Cost", "QALY", "NMB"),
  strategies = c("SoC", "Anifrolumab + SoC"),  # Only SOC and Anifrolumab since Belimumab + SOC dominated
  progress = TRUE
)

# Create tornado plots
# Select the net monetary benefit (NMB) owsa object
my_owsa_NMB <- l_owsa_det$owsa_NMB

summary(my_owsa_NMB)

# Visualize as tornado diagram
owsa_tornado_plot <- owsa_tornado(my_owsa_NMB)
print(owsa_tornado_plot)

# Plot outcomes of each strategy over each parameter range
owsa_strategy_plot <- plot(my_owsa_NMB, n_x_ticks = 5)
print(owsa_strategy_plot)

# Visualize optimal strategy (max NMB) over each parameter range
owsa_optimal_plot <- owsa_opt_strat(my_owsa_NMB)
print(owsa_optimal_plot)

# Save plots
#ggsave("figures/owsa_tornado_plot.png", plot = owsa_tornado_plot, width = 10, height = 8, dpi = 300)
#ggsave("figures/owsa_strategy_plot.png", plot = owsa_strategy_plot, width = 12, height = 10, dpi = 300)
#ggsave("figures/owsa_optimal_plot.png", plot = owsa_optimal_plot, width = 10, height = 8, dpi = 300)
```

##Cost Threshold Analysis for Anifrolumab

```{r}
cost_range <- seq(10000, 23000, by = 1000)

#Store results
icer_df <- data.frame(
  cost_anifrolumab = numeric(),
  ICER = numeric()
)

# Loop through each cost
for (cost_val in cost_range) {
  params <- list(cost_anifrolumab = cost_val)
  
  results <- run_lupus_model(params)
  
  # Extract cost and QALY for each strategy
  soc <- results[results$Strategy == "SoC", ]
  anif <- results[results$Strategy == "Anifrolumab + SoC", ]
  
  # Calculate ICER
  delta_cost <- anif$Cost - soc$Cost
  delta_qaly <- anif$QALY - soc$QALY
  
  icer <- ifelse(delta_qaly == 0, NA, delta_cost / delta_qaly)
  
  # Save results
  icer_df <- rbind(icer_df, data.frame(
    cost_anifrolumab = cost_val,
    ICER = icer
  ))
}

library(ggplot2)

ggplot(icer_df, aes(x = cost_anifrolumab, y = ICER)) +
  geom_line(color = "#15979d", size = 1.2) +
  geom_hline(yintercept = 150000, linetype = "dashed", color = "black") +  # WTP threshold
  annotate("text", min(cost_range) + 1000, y = 150000 + 5000,
           label = "WTP Threshold: $150,000/QALY", 
           hjust = 0, vjust = 0, fontface = "bold") +
  labs(
    title = "ICER vs. Cost of Anifrolumab",
    x = "6-Month Cost of Anifrolumab ($)",
    y = "ICER ($/QALY)",
    labels=scales::label_dollar(),
    plot.title = element_text(hjust = 0.5, face = "bold")
  ) +
  theme_minimal()
```

